---
title: "Women's World Cup"
author: "Joshua Ganz, Julian Mucha, Eric Rwabuhihi"
date: "October 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
#install packages flexdashboard, tidyverse, ggimage, ggthemes, plotly, cowplot, sf, rworldmap, and rgeos and run library.
if (!require("pacman")) install.packages("pacman")
pacman::p_load(flexdashboard, tidyverse, ggimage, ggthemes, plotly, cowplot, sf, rworldmap, rgeos, countrycode)
```

This is the code to run Amanda Peterson Plunkett's Women's World Cup #tidytuesday submission.  Her geom_image was excluded from her code since it is a picture of a soccer ball saved to her local drive.
```{r message=F, warning=F}

#Get the data
wwc_outcomes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv")
squads <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
codes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")

#get alternative country codes
more_codes <- countrycode::codelist_panel %>% 
  group_by(country.name.en) %>% 
  top_n(1, year) %>%
  select(country.name.en, ioc, iso2c, iso3c, genc3c, fips)

#join data with codes
wwc_outcomes_wcodes <- dplyr::left_join(wwc_outcomes, codes, by = "team") %>%
  select(year, team, score, country) %>%
  left_join(more_codes, by = c("team" = "ioc")) %>%
  mutate(year = as.character(year))

#Fix England:
wwc_outcomes_wcodes[wwc_outcomes_wcodes$country == "England", "iso2c"] <- "GB"

#Limit plot to top scoring countries 
top_countries <- wwc_outcomes %>%
  group_by(team) %>%
  summarize(total_score = sum(score)) %>%
  top_n(11, total_score) 

#plot
ggplot(inner_join(wwc_outcomes_wcodes, top_countries), 
       aes(reorder(country, total_score), score, fill = year)) + 
  geom_col(position = position_stack(reverse = TRUE)) + #adjusted position_stack to have years increase from left to right
  ggthemes::scale_fill_tableau(name="Year") +           #nice colors
  coord_flip() +
  labs(
    title = "Womens World Cup Soccer: Total Goals (1991 - 2019)",
    caption = "Data from https://data.world/sportsvizsunday/womens-world-cup-data"
  ) +
  geom_flag(y = -5, aes(image = iso2c)) +
  expand_limits(y = -5) +
 # geom_image(aes(x = "France", y = 120, 
                 #image = "~/tidytuesday/2019/week28/soccer2.png"),
            # size = 0.15) +
  theme_minimal() +
  theme(title = element_text(size=14),
        panel.grid = element_blank(), 
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_text(angle = 180)) + 
  guides(fill = guide_legend(reverse = TRUE))  #put 2019 at top of legend

```


The code below is taking the same data but displaying it in a different way (Modified Selection)
```{r message=F, warning=F}
#Get the data
critique_wwc_outcomes <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/wwc_outcomes.csv")
critique_squads <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/squads.csv")
critique_codes <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-09/codes.csv")

#join data with codes
critique_wwc_outcomes_wcodes <- left_join(critique_wwc_outcomes, codes, by = "team") %>%
  mutate(year = as.character(year))%>%
  select(year, country, score)

#Limit plot to top scoring countries 
critique_top_countries <- critique_wwc_outcomes_wcodes %>%
  group_by(country) %>%
  summarize(total_score = sum(score)) %>%
  top_n(9, total_score) 

#add rows for years countries were not in the WC
newrow = c("1991", "England", ' ')
critique_wwc_outcomes_wcodes = rbind(critique_wwc_outcomes_wcodes,newrow)
newrow1 = c("1999", "England", ' ')
critique_wwc_outcomes_wcodes = rbind(critique_wwc_outcomes_wcodes,newrow1)
newrow2 = c("2003", "England", ' ')
critique_wwc_outcomes_wcodes = rbind(critique_wwc_outcomes_wcodes,newrow2)
newrow3 = c("2011", "China PR", ' ')
critique_wwc_outcomes_wcodes = rbind(critique_wwc_outcomes_wcodes,newrow3)
newrow4 = c("1991", "Australia", ' ')
critique_wwc_outcomes_wcodes = rbind(critique_wwc_outcomes_wcodes,newrow4)

#Change score 'type' to numeric
critique_wwc_outcomes_wcodes$score = as.numeric(critique_wwc_outcomes_wcodes$score)

#join data
critique_joined_wwc <- inner_join(critique_wwc_outcomes_wcodes, critique_top_countries)
critique_joined_wwc$country <- gsub("China PR", "China", critique_joined_wwc$country)

#create data to plot total tournament goals by country and year
critique_plot <- critique_joined_wwc %>%
  group_by(year, country)%>%
  summarize(year_score=sum(score))#%>%

#create data for total goals over all years
critique_plot1 <- critique_joined_wwc %>%
  group_by(year, country)%>%
  summarize(total_score = median(total_score))

#merge plot and plot1 to use for line graph
critique_plot_final <- merge(critique_plot, critique_plot1, by = c("year", "country"))
colnames(critique_plot_final)[4] <- "total_goals_combine_all"
colnames(critique_plot_final)[3] <- "year_goals"

#plot goals scored by year and make it interactive
critique_goal_plot <- ggplot(critique_plot_final, mapping = aes(year, year_goals, color = country, total_goals_combine_all= total_goals_combine_all)) +
  geom_point(mapping = aes(group = country), size=2) +
  geom_path(mapping = aes(group=country), na.rm = FALSE) +
  facet_wrap(~country)+
  labs(y = "", x = "",
    title = "Women's World Cup Goals per Year (1991 - 2019: Top 9 Countries Overall)",
    caption = "Data from https://data.world/sportsvizsunday/womens-world-cup-data"
  ) +
  theme_bw() +
  theme(panel.background = element_blank(),
           panel.grid.major.x = element_blank(),
           legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete()+
  scale_y_continuous(breaks = c(0,4,8,12,16,20,24,28,32), limits = c(0,32))
 
ggplotly(critique_goal_plot, tooltip = (c("year", "year_goals", "total_goals_combine_all")))
```

This code is the beginning of the group's ideas for a future visualization to incorporate more of a geographical theme, eventually including dropdown menus, cumulative goal data each latter year selected, etc.
```{r, message=F, warning=F}

country_lats_longs <- readr::read_csv("https://raw.githubusercontent.com/albertyw/avenews/master/old/data/average-latitude-longitude-countries.csv")
wwc_outcomes_wcodes1 <- wwc_outcomes_wcodes %>%
  mutate(Country = country.name.en)

#creates total goals scored by each team by year
scores_by_team <- wwc_outcomes %>%
  group_by(year, team) %>%
  summarise(total_score = sum(score))

scores_by_year <- right_join(scores_by_team, wwc_outcomes_wcodes1, by=c('team' = 'team')) %>%
  select(team, year.x, total_score, country.name.en, Country)

scores_by_year$Country[is.na(scores_by_year$Country)] <- 'Nigeria'

wwc_outcomes_wcodes1 <- wwc_outcomes_wcodes1 %>%
  group_by(year, team, country, country.name.en, iso2c, iso3c, genc3c, fips, Country) %>%
  summarise(n = n()) %>%
  mutate(total_score = sum(wwc_outcomes$score))

wwc_outcomes_wlatsandlongs <- left_join(scores_by_year, country_lats_longs, by =c('Country' = 'Country'))

map.world <- map_data('world')
#View(map.world)
wwc_outcomes_wlatsandlongs <- wwc_outcomes_wlatsandlongs %>%  mutate(Country = recode(Country, `United States` = 'USA'
                                 , `United Kingdom` = 'UK'
                                 , `Congo, Democratic Republic of the` = 'Democratic Republic of the Congo'
                                 , `Trinidad and Tobago` = 'Trinidad'
                                 , `Congo, Republic of the` = 'Republic of Congo'
                                 )
                   )

#filters the source dataset for only data relating to 1991
wwc_outcomes_wlatsandlongs1 <- wwc_outcomes_wlatsandlongs %>%
  filter(year.x == 1991)

#anti_join(wwc_outcomes_wlatsandlongs1, map.world, by = c('Country' = 'region'))

#joins the latitude and longitude data with the source data so it can be mapped
map.goals <- left_join( map.world, wwc_outcomes_wlatsandlongs1, by = c('region' = 'Country'))

#builds the map
ggplot(map.goals, aes( x = long, y = lat, group = group)) +
  #the country fill is based on the total goals scored for each of the world cup years
  geom_polygon(aes(fill = total_score)) +
  #creates the fill gradient for the countries based on total score
  scale_fill_gradient(low = "#56B1F7", high = "#132B43") +
  #creates the Title, subtitle, and caption of the graph
  labs(fill = 'Goals'
       ,title = 'Goals Scored per Country'
       ,subtitle = '1991'
       ,caption = "Source: https://www.sharpsightlabs.com/blog/highlight-countries-on-map/"
       ,x = NULL
       ,y = NULL) +
#deals with the aesthetics of the rest of the graph including text color, background color, legend color, etc.
  theme(text = element_text(family = 'Gill Sans', color = '#000000')
        ,plot.title = element_text(size = 28)
        ,plot.subtitle = element_text(size = 14)
        ,axis.ticks = element_blank()
        ,axis.text = element_blank()
        ,panel.grid = element_blank()
        ,panel.background = element_rect(fill = '#bbd4dd')
        ,plot.background = element_rect(fill = '#333333')
        ,legend.position = c(.18,.36)
        ,legend.background = element_blank()
        ,legend.key = element_blank()
        ) +
#adjusts the alignment, font color, and face type of the title, subtitle, and caption
  theme(plot.title = element_text(hjust=0.5, color = 'white'), plot.subtitle = element_text(hjust = 0.5, size = 15, face = 'bold', color = 'white'), plot.caption = element_text(size = 8, color = 'white'))
```